Pyro-Velocity: Probabilistic RNA Velocity inference from single-cell data

### 2024-08-19 reproducibility/manuscript/manuscript.qmd
docs(manuscript): specify newpage breaks
Signed-off-by: Cameron Smith <cameron.ray.smith@gmail.com>

---
title: "Pyro-Velocity: Probabilistic RNA Velocity inference from single-cell data"
format:
  nature-pdf:
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-nature
    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
    # to use the bst styles from the upstream template.
    cite-method: citeproc
    keep-tex: true
    equal-margins: false
  html:
    code-fold: true
    html-math-method: mathjax
  docx: default
csl: _extensions/sciexp/nature/csl/springer-basic-brackets-no-et-al.csl
author:
  - name: Qian Qin
    affiliations:
      - name: Massachusetts General Hospital Research Institute, Charlestown, MA
        id: 1
        department: Molecular Pathology Unit
        city: Charlestown
        state: MA
        postal-code: 02129
      - name: Harvard Medical School, Charlestown, MA
        id: 2
        department: Massachusetts General Hospital Cancer Center
        city: Charlestown
        state: MA
        postal-code: 02129
      - name: Massachusetts General Hospital, Boston, MA
        id: 3
        department: Center for Regenerative Medicine
        city: Boston
        state: MA
        postal-code: 02114
      - name: Harvard University, Cambridge, MA
        id: 4
        department: Harvard Stem Cell Institute
        city: Cambridge
        state: MA
        postal-code: 02139
      - name: Cambridge, MA
        id: 5
        department: Broad Institute of MIT and Harvard
        city: Cambridge
        state: MA
        postal-code: 02142
    # attributes:
    #   equal-contributor: true
  - name: Eli Bingham
    affiliations:
      - ref: 5
      - name: Cambridge, MA
        id: 7
        department: Basis Research Institute
        city: Cambridge
        state: MA
        postal-code: 02142
  - name: Gioele La Manno
    affiliations:
      - name: Laboratory of Neurodevelopmental Systems Biology, Lausanne, Switzerland
        id: 6
        department: Brain and Mind Institute, School of Life Sciences, École Polytechnique Fédérale de Lausanne (EPFL)
        city: Lausanne
        state: Switzerland
        postal-code: 1015
    email: gioele.lamanno@epfl.ch
    attributes:
      corresponding: true
  - name: David M. Langenau
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
      - ref: 4
    email: dlangenau@mgh.harvard.edu
    attributes:
      corresponding: true
  - name: Luca Pinello
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 5
    email: lpinello@mgh.harvard.edu
    attributes:
      corresponding: true
abstract: |
  Single-cell RNA Velocity has dramatically advanced our ability to model cellular differentiation and cell fate decisions. However, current preprocessing choices and model assumptions often lead to errors in assigning developmental trajectories. Here, we develop, Pyro-Velocity, a Bayesian, generative, and multivariate RNA Velocity model to estimate the uncertainty of cell future states. This approach models raw sequencing counts with the synchronized cell time across all expressed genes to provide quantifiable and improved information on cell fate choices and developmental trajectory dynamics.
keywords: [RNA velocity, single-cell RNA sequencing, probabilistic generative modeling, cell fate, developmental trajectory inference]
bibliography: bibliography.bib
crossref:
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Figure S
      space-before-numbering: false
      latex-list-of-description: Supplementary Figure
---

@@

@@

@@

@@

\begin{table}[h]
\caption{LaTeX caption text}\label{tbl-data}%
\begin{tabular}{@{}llll@{}}
\toprule
Column 1 & Column 2  & Column 3 & Column 4\\
\midrule
row 1    & data 1   & data 2  & data 3  \\
row 2    & data 4   & data 5\footnotemark[1]  & data 6  \\
row 3    & data 7   & data 8  & data 9\footnotemark[2]  \\
\botrule
\end{tabular}
\footnotetext{Source: This is an example of table footnote. This is an example of table footnote.}
\footnotetext[1]{Example for a first table footnote. This is an example of table footnote.}
\footnotetext[2]{Example for a second table footnote. This is an example of table footnote.}
\end{table}

# References
<!-- the pattern below controls the placement of the references -->
:::{#refs}
:::

>>>>

---
title: "Pyro-Velocity: Probabilistic RNA Velocity inference from single-cell data"
format:
  nature-pdf:
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-nature
    classoption: 
      - lineno
    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
    # to use the bst styles from the upstream template.
    cite-method: citeproc
    keep-tex: true
    equal-margins: false
    include-in-header: header.tex
  html:
    code-fold: true
    html-math-method: mathjax
  docx: default
csl: _extensions/sciexp/nature/csl/springer-basic-brackets-no-et-al.csl
author:
  - name: Qian Qin
    affiliations:
      - name: Massachusetts General Hospital Research Institute, Charlestown, MA
        id: 1
        department: Molecular Pathology Unit
        city: Charlestown
        state: MA
        postal-code: 02129
      - name: Harvard Medical School, Charlestown, MA
        id: 2
        department: Massachusetts General Hospital Cancer Center
        city: Charlestown
        state: MA
        postal-code: 02129
      - name: Massachusetts General Hospital, Boston, MA
        id: 3
        department: Center for Regenerative Medicine
        city: Boston
        state: MA
        postal-code: 02114
      - name: Harvard University, Cambridge, MA
        id: 4
        department: Harvard Stem Cell Institute
        city: Cambridge
        state: MA
        postal-code: 02139
      - name: Cambridge, MA
        id: 5
        department: Broad Institute of MIT and Harvard
        city: Cambridge
        state: MA
        postal-code: 02142
    # attributes:
    #   equal-contributor: true
  - name: Eli Bingham
    affiliations:
      - ref: 5
      - name: Cambridge, MA
        id: 7
        department: Basis Research Institute
        city: Cambridge
        state: MA
        postal-code: 02142
  - name: Gioele La Manno
    affiliations:
      - name: Laboratory of Neurodevelopmental Systems Biology, Lausanne, Switzerland
        id: 6
        department: Brain and Mind Institute, School of Life Sciences, École Polytechnique Fédérale de Lausanne (EPFL)
        city: Lausanne
        state: Switzerland
        postal-code: 1015
    email: gioele.lamanno@epfl.ch
    attributes:
      corresponding: true
  - name: David M. Langenau
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 3
      - ref: 4
    email: dlangenau@mgh.harvard.edu
    attributes:
      corresponding: true
  - name: Luca Pinello
    affiliations:
      - ref: 1
      - ref: 2
      - ref: 5
    email: lpinello@mgh.harvard.edu
    attributes:
      corresponding: true
abstract: |
  Single-cell RNA Velocity has dramatically advanced our ability to model cellular differentiation and cell fate decisions. However, current preprocessing choices and model assumptions often lead to errors in assigning developmental trajectories. Here, we develop, Pyro-Velocity, a Bayesian, generative, and multivariate RNA Velocity model to estimate the uncertainty of cell future states. This approach models raw sequencing counts with the synchronized cell time across all expressed genes to provide quantifiable and improved information on cell fate choices and developmental trajectory dynamics.
keywords: [RNA velocity, single-cell RNA sequencing, probabilistic generative modeling, cell fate, developmental trajectory inference]
bibliography: bibliography.bib
crossref:
  custom:
    - kind: float
      key: suppfig
      latex-env: suppfig
      reference-prefix: Figure S
      space-before-numbering: false
      latex-list-of-description: Supplementary Figure
---

\newpage

@@

\newpage

# References
:::{#refs}
:::

\newpage

@@

\newpage

@@

![Extended version of @fig-model-results for evaluating Pyro-Velocity (Model 1) and scVelo on the PBMC and pancreas datasets. a. Left figure shows the scVelo vector field using the default user options, right figure shows the evaluation of the scVelo vector field using different combinations of user options within Epsilon cells. b. Pyro-Velocity additional single cell level uncertainty diagnostic plots on the fully mature PBMC dataset, from left to right: 1. standard deviation of posterior samples of shared time, 2. Rayleigh test of posterior samples of the vector field (the title reports false discovery rate after p-value correction for multiple tests across cells), 3. transcriptional state uncertainty of predicted raw read count per cell. The contour lines show the top 10% of the most uncertain cells. c. For the pancreas dataset, the figures show from left to right: UMAP rendering of 1. Cytotrace, 2. scVelo latent time, 3. Posterior mean of Pyro-Velocity shared time averaged across 30 posterior samples, 4. Top 50 markers overlapping between Pyro-Velocity and scVelo (Methods, Supplementary Table 1-2), 5. The same uncertainty diagnostic plots per cell as in b. d. The same plot as Fig. 2d. e. Selection of the top negatively correlated genes with the cell shared time based on the posterior mean on denoised spliced expression and negative mean absolute error of gene expression prediction (Methods). f. Phase portraits, rainbow plots, and UMAP rendering of posterior mean of gene expression for the genes selected in e.](figures/fig.pdf){#suppfig-model-results}

@@

<!-- \begin{table}[h]
\caption{LaTeX caption text}\label{tbl-data}%
\begin{tabular}{@{}llll@{}}
\toprule
Column 1 & Column 2  & Column 3 & Column 4\\
\midrule
row 1    & data 1   & data 2  & data 3  \\
row 2    & data 4   & data 5\footnotemark[1]  & data 6  \\
row 3    & data 7   & data 8  & data 9\footnotemark[2]  \\
\botrule
\end{tabular}
\footnotetext{Source: This is an example of table footnote. This is an example of table footnote.}
\footnotetext[1]{Example for a first table footnote. This is an example of table footnote.}
\footnotetext[2]{Example for a second table footnote. This is an example of table footnote.}
\end{table} -->

\newpage
